???A quantitative framework for investigating the reliability of network construction

### 2017-09-17 manuscript/CI_interactions.tex
upscaling uncertainty

\author{Dominique Gravel, \& Alyssa R. Cirtwill$^{1}$, Anna Ekl\"{o}f, Tomas Roslin, Kate Wootton}
\date{\small$^1$Department of Physics, Chemistry\\ 
and Biology (IFM)\\ 
Link\"{o}ping University\\
Link\"{o}ping, Sweden\\
% \medskip
% $^\dagger$ Corresponding author:\\
% alyssa.cirtwill@gmail.com\\
% +46 723 158464\\
 }

@@

@@

      prior_web_data=read.csv(path_to_Dryad_file)
      # Remove extra data on parasitoids, etc.
      prior_web_data[,31:57]<-NULL
      prior_web_data[,4:26]<-NULL
      prior_web_data[,2]<-NULL
      # Build a web with interactions=1 for any galler observed on any genotype
      prior_web=matrix(nrow=26,ncol=4)
      for(r in 1:length(levels(prior_web_data$Genotype))){
        gen=levels(prior_web_data$Genotype)[r]
        subset=prior_web_data[which(prior_web_data$Genotype==gen),]
        for(c in 1:4){
          if(sum(subset[,2+c])>0){
            prior_web[r,c]=1
          } else {
            prior_web[r,c]=0
          }
        }        
      }
      # Now get the degree distributions from the web
      deg_dist_Salix=rowSums(prior_web)/ncol(prior_web)
      # Remove one genotype that never interacted
      deg_dist_Salix<-deg_dist_Salix[which(rowSums(prior_web)>0)]
      deg_dist_galler=colSums(prior_web)/nrow(prior_web)
      # Interaction probabilities are the product of plant and galler probabilities
      int_probs=as.numeric(deg_dist_galler%*%t(deg_dist_Salix))

      Using this degree distribution we can obtain prior values for $\alpha$ and $\beta$ using the code above. We obtain
      $\alpha$=2.807855, $\beta$=2.342815.

      For species where no co-occurances were observed, we can calculate the MLE estimates for the mean and variance of $\omega_{ij}$ directly from this prior. These are:
      $\hat{\omega_{ij}}$=0.5451437, var($\omega$)=0.0481417.
      This is consistent with the prior data, where the average interaction probability based on the degree distributions of Salix genotypes and gallers was 0.541. It may not be consistent with the degree distributions in our current dataset, however, especially since we are concerned with Salix species rather than genotypes within a species. We therefore want to update our distribution using known data from known co-occurances between species.

@@

      $\hat{\omega_{ij}}$=7.405645e-03, var($\omega$)=1.938755e-05.

      Realistic values for n when k=0, based on a sample of 20 random sites:
      n=1: 0.4565121 0.0403385
      n=2: 0.39267022 0.03335076
      n=3: 0.34449379 0.02770543
      n=4: 0.30684696 0.02324331

@@

  [[What if we have only one site? Can we use the number of observations of a pair of species interacting with any partners as the number of times they co-occur, and the number of interactions within the pair as k? This will be sort-of valid for insects but not for plants when researchers camp out by plant individuals... ]]

>>>>

\author{Dominique Gravel, D. Carstensen, Timoth\'{e}e Poisot, Daniel B. Stouffer, \& Alyssa R. Cirtwill$^{1}$, others??}
\date{\small$^1$Department of Physics, Chemistry\\ 
and Biology (IFM)\\ 
Link\"{o}ping University\\
Link\"{o}ping, Sweden\\
% \medskip
% $^\dagger$ Corresponding author:\\
% alyssa.cirtwill@gmail.com\\
% +46 723 158464\\
 }

@@

It is fairly straigthforward to compute most of network metrics when the different $\lambda$ of the adjacency matrix are known, without variance (Poisot et al. 2015). Several of these metrics derive directly from quantitative indices of network structure. It is not as easy however to understand how the uncertainty in these estimates do influence network metrics. There are multiple non-linearities in the computation of these metrics and that, because of Jensen's inequality (the principle stating that the average of non-linear function of a stochastic variable differs from the function of the average of that variable), any uncertainty could bias the average estimate of the metric and influence its variance as well. For instance, among all of the pairs of species for which only few co-occurrences are observed, many of them are truly interacting. If we consider all of these as having no links, then correcting for these false absence would inflate the number of links and therefore influence metrics such as connectance, nestedness and degree distribution. Computing network properties from probabilistic networks is one thing, but then computing them with uncertainties in these probabilities is another problem, far more challenging. Unfortunately, for now we could only rely on computer simulations to get an answer to this problem. In the following section, we will provide an example with the Salix dataset and show how the uncertainty arising from both absences of co-occurrences and false absences of interactions impact these metrics. We find uncertainty influences both the average and the variance of network metrics, putting in perspective traditional network comparison studies where these metrics are assumed fixed entities.

@@

      prior_web=read.csv(path_to_Dryad_file)
      deg_dist_Salix=with(prior_web,tapply(gall_total_rich,Genotype,mean))/4
      gallers=cbind(prior_web$Genotype,prior_web[,27:30])
      aSG_means=with(prior_web,tapply(aSG_abund,Genotype,mean))
      rG_means=with(prior_web,tapply(rG_abund,Genotype,mean))
      vLG_means=with(prior_web,tapply(vLG_abund,Genotype,mean))
      SG_means=with(prior_web,tapply(SG_abund,Genotype,mean))
      deg_dist_galler=c(length(which(aSG_means>0)),length(which(rG_means>0)),length(which(SG_means>0)),length(which(rG_means>0))      )/26
      int_probs=as.numeric(deg_dist_galler\%*\%t(deg_dist_Salix))

      pars=fitdistr(x=int_probs,"beta",start=list(shape1=1,shape2=1))$estimate
      alpha=pars[[1]]
      beta=pars[[2]]
      mu=alpha/(alpha+beta)
      num=alpha*beta
      den=(alpha+beta)**2*(alpha+beta)
      sigma2=num/den

      Using this degree distribution we can obtain prior values for $\alpha$ and $\beta$ using the code above. We obtain
      $\alpha$=2.175656, $\beta$=5.361083.

      For species where no co-occurances were observed, we can calculate the MLE estimates for the mean and variance of $\omega_{ij}$ directly from this prior. These are:
      $\hat{\omega_{ij}}$=0.2886734, var($\omega$)=0.02724535.

@@

      $\hat{\omega_{ij}}$=0.005702349, var($\omega$)=1.486051e-05.

      Realistic values for n when k=0, based on a sample of 20 random sites:
      n=1: 0.25485794 0.02224566
      n=2: 0.22813413 0.01846427
      n=3:
      n=4:

@@
